\LARGE \bf Imposing a Weight Norm Constraint for Neuro–Adaptive Control

### 2025-03-19 manuscript.tex
subtitute notations using template

@@

@@

\subsection{Notation}
In this study, the following notation is used.

\begin{itemize}
    \item $\otimes$ denotes the Kronecker product \cite{Bernstein:2009aa}.
    \item $\mv x_{(i)}$ denotes the $i\textsuperscript{th}$ element of vector $\mv x$.    
    % \item $A_{(i,j)}$ is the $i\textsuperscript{th}$ row-$j\textsuperscript{th}$ column element of matrix $A$.
    \item $\text{row}_j(\mm A)$ denotes the $j\textsuperscript{th}$ row of $\mm A\in\mathbb{R}^{n\times m}$. 
    % \item $\text{col}_i(A)$ and $\text{row}_j(A)$ denote the $i\textsuperscript{th}$ column and $j\textsuperscript{th}$ row of $A\in\mathbb{R}^{n\times m}$, respectively.
    \item $\text{vec}(\mm A)\triangleq [\text{row}_1(\mm A^\top )  ,\cdots,\text{row}_m(\mm A^\top )]^\top $ for $\mm A\in\mathbb{R}^{n\times m}$.
    % \item $\text{vec}(A)\triangleq [\text{col}_1(A)^\top  ,\cdots,\text{col}_m(A)^\top  ]^\top   $ for $A\in\mathbb{R}^{n\times m}$.
    \item $\lambda_\text{min}(A)$ denotes the minimum eigenvalue of $A\in\mathbb{R}^{n\times n}$.
\end{itemize}

@@

Consider an uncertain Euler‒Lagrange system modeled as
\begin{equation}
    \mm M(\mv q)\ddot {\mv q} + \mm C(\mv q,\dot {\mv q})\dot {\mv q} + \mm G(\mv q) + \mm F(\mv q) = \mv \tau
    \label{eq. system dynamics 1}
\end{equation}
where $\mv q\in \mathbb{R}^n$ and $\mv \tau\in\R^n$ denote the generalized coordinate and the control input, respectively; $\mm M(\mv q)\in\mathbb{R}^{n\times n}$, $\mm C(\mv q,\dot {\mv q})\in\mathbb{R}^{n\times n}$, and $\mm G(\mv q)\in\mathbb{R}^{n}$ denote the unknown system function matrices; and $\mm F(\mv q)\in\mathbb{R}^{n}$ denotes the external force.
Using the user-designed matrices $\mm M_0>0,\mm C_0$ and $\mm G_0$, \eqref{eq. system dynamics 1} can be represented as 
\begin{equation}
    \mm M_0\ddot {\mv q}+\mm C_0\dot {\mv q}+\mm G_0 = \mv \tau + \mv f(\mv q,\dot {\mv q},\ddot {\mv q})
    \label{eq. system dynamics 2}
\end{equation}
where $\mv f(\mv q,\dot {\mv q},\ddot {\mv q}) \triangleq -(\mm M-\mm M_0)\ddot {\mv q}-(\mm C-\mm C_0)(\mv q,\dot {\mv q})\dot {\mv q} -(\mm G-\mm G_0) -\mm F(\mv q)$ denotes the residual unknown term.

Hence, the objective of the control design is to make $\mv q$ track the continuously differentiable desired trajectory $\mv q_d(t):\R\to \R^n$ under the unknown terms $\mv f$.

@@

The backstepping control-based approach is utilized to generate a reference signal $\mv z^*\triangleq -{k_q}\tilde {\mv q}+\dot {\mv q}_d$ for $\mv {z}\triangleq \dot {\mv q}$, where $\tilde {\mv q}\triangleq \mv {q} - {\mv q}_d$ and $k_q\in\R_{>0}$.
The desired stabilizing controller can be designed as follows:
\begin{equation}
    \mv \tau^* = 
    -\mm M_0\cdot ({k_z}\tilde {\mv z})
    + 
        ( 
            -\mm M_0\tilde {\mv q}+\mm C_0{\mv z}+\mm G_0-\mv f+\mm M_0 \dot {\mv z}^*
        )
    \label{eq. desired control}
\end{equation}
where $\tilde {\mv z}\triangleq \mv z-\mv z^*$ and $k_z\in\R_{>0}$.
Note that the desired controller cannot be realized because of $\mv f$.

To approximate the desired controller, an NN is utilized.
The NN with a single hidden layer is represented as 
\begin{equation}
    \mv \Phi({\mv q_{NN}};\mv \theta)\triangleq \mm V_1^\top \mv \phi(\mm V_0^\top {\mv q_{NN}})
\end{equation}
where ${\mv q_{NN}}\in\R^{l_0+1}$ denotes the NN input vector, $\mm V_i\in\R^{(l_i+1)\times l_{i+1}},\ i\in[0,1]$ denotes the weight matrix of the $i\textsuperscript{th}$ layer and $\mv \phi:\R^{l_1}\to\R^{l_1+1}$ denotes the activation function layer.
The activation function layer consists of an elementwise nonlinear function $\sigma(\cdot)$ and an augmented $1$ to combine the bias terms in the weight matrix (\ie $\mv \phi(\mv x) = [\sigma(\mv x_{(1)}),\cdots, \sigma(\mv x_{(l_1)}), 1]^\top $).
For further simplicity, let $\mv \theta\triangleq[\mv \theta_1^\top ,\mv \theta_0^\top ]^\top \in\R^{\Xi}$ denote the total weight vector, where $\mv \theta_i\triangleq \text{vec}(\mm V_i)\in\R^{\Xi_i}$ denotes the vectorized weights.
$\Xi_i=(l_i+1)\cdot l_{i+1}$ and $\Xi=\Xi_0+\Xi_1$ denote the number of each layer and total weights, respectively.

Using this NN, the desired controller $\mv \tau^*$ can be approximated by the ideal weight vector $\mv \theta^*$ for a compact subset $\Omega_{NN}\in\R^{l_0+1}$ to $\epsilon$ accuracy such that $\sup_{{\mv q_{NN}}\in\Omega_{NN}} \allowbreak \Vert\mv \Phi({\mv q_{NN}};\mv \theta^*) - \mv \tau^*\Vert=\epsilon<\infty$ \cite{Cybenko:1989aa}.
The ideal weight vector $\mv \theta^*$ is typically assumed to be bounded.
Then, using the estimated weight vector $\hat{\mv \theta}=[\hat{\mv \theta}_1^\top ,\hat{\mv \theta}_0^\top ]^\top $ of $\mv \theta^*=[\mv \theta_1^{*\top},\mv \theta_0^{*\top}]^\top $, the desired controller $\mv \tau^*\approx -\mv \Phi(\mv q_{NN};\mv \theta^*)-\epsilon$ can be approximated as follows:
\begin{equation}
    \mv \tau \triangleq -\mv \Phi({\mv q_{NN}};\hat{\mv \theta}).
    \label{eq. control law}
\end{equation}
For further sections, let $\mv \Phi^*\triangleq\mv \Phi({\mv q_{NN}};\mv \theta^*)$ and $\mv \phi^* \triangleq\mv \phi(\mm V_0^{*\top}{\mv q_{NN}})$, and let $\hat{\mv \Phi}\triangleq{\mv \Phi}({\mv q_{NN}};\hat{\mv \theta})$, $\hat{\mv \phi} \triangleq\mv \phi(\hat {\mm V}_0^{\top}{\mv q_{NN}})$ and $\hat{\mv \phi}' \triangleq \partial \hat{\mv \phi}/\partial (\hat {\mm V}_0^\top \mv q_{NN})$.

Using \eqref{eq. system dynamics 2}, \eqref{eq. desired control}, and \eqref{eq. control law}, the error dynamics can be obtained as the first-order system of augmented error $\mv \xi\triangleq[\tilde {\mv q}^\top ,\tilde {\mv z}^\top ]^\top \in\R^{2n}$ as follows:
\begin{equation}
    \dot{\mv \xi} 
    = 
    \mm A_\xi{\mv \xi}+\mm B_\xi(\mv \Phi^*-\hat{\mv \Phi}+\epsilon)
\end{equation}
where 
\begin{equation}
    \mm A_\xi \triangleq 
    \begin{bmatrix}
        -{k_q} \mm I_{n\times n} &\mm I_{n\times n}
        \\
        -\mm I_{n\times n}& -{k_z} \mm I_{n\times n}
    \end{bmatrix}
    ,\ 
    \mm B_\xi \triangleq 
    \begin{bmatrix}
        \mm 0_{n\times n}\\\mm M_0^{-1}
    \end{bmatrix}.
\end{equation}

@@

Using the augmented error $\mv \xi$, the quadratic objective function is defined as
\begin{equation}
    J(\mv \xi;\hat{\mv \theta})\triangleq
    {1\over 2}
    \mv \xi^\top \mm W\mv \xi
\end{equation}
where $\mm W=\mm W^\top >0$ denotes the weighting matrix.

For the boundedness of the weights, weight norm constraints are imposed on the adaptation process such that $c_{\theta_i}=\Vert{\hat{\mv\theta}}_i\Vert^2-\bar\theta^2_i,\ i\in[0,1]$.
% Then the objective of the adaptation process can be reformulated into the constrained optimization problem as follows:
The control problem can be reformulated into a constrained optimization problem as follows:
\begin{equation}
    \begin{matrix}
        \min_{{\hat{\mv\theta}}} \ J(\mv \xi;{\hat{\mv\theta}})
        \\ \\
        \begin{aligned}
        \text{subject to\;}&c_{j}\le0, \quad j\in\mathcal I = [\theta_0,\theta_1].
        \end{aligned}
    \end{matrix}
\end{equation}
In this optimization problem, $\xi$ is considered a predefined parameter.
The corresponding Lagrangian function is defined as
\begin{equation}
    L(
        \mv \xi,
        {\hat{\mv\theta}},
        [\lambda_j]_{j\in\mathcal I}
    )
    \triangleq 
    J(\mv \xi;{\hat{\mv\theta}})
    +
    \sum_{j\in\mathcal I} \lambda_jc_j({\hat{\mv\theta}})
    .
\end{equation}
% where $\mathcal I=\{j\ \vert \ c_j\ge 0\}$ denotes the active set.
To solve the dual problem $\min_{{\hat{\mv\theta}}} \max_{[\lambda_j]_{j\in\mathcal I}}  L$, the adaptation law is derived as follows:
\begin{align}
        \dot {{\hat{\mv\theta}}}&=-\alpha {\partial L\over\partial {\hat{\mv\theta}}}
        =-\alpha 
        \bigg(
        {\partial J\over \partial {\hat{\mv\theta}}}+\sum_{j\in\mathcal{A}}
        \lambda_j {\partial c_j\over \partial {\hat{\mv\theta}}}
        \bigg),
    \label{eq. adaptation law th}
        \\
        \dot\lambda_j& = \beta_j{\partial L\over\partial \lambda_j} = \beta_j c_j ,
        \quad\quad\quad\quad      \      
        \forall j\in\mathcal I,
    \label{eq. adaptation law L}
        \\
        \lambda_j & = \max(\lambda_j,0) ,
        \quad\quad\quad\quad\ \ \ \ \ 
        \forall j\in\mathcal I,
    \label{eq. adaptation law L max}
\end{align}
where $\alpha\in\R_{>0}$ and $\beta_j\in\R_{>0}$ denote the adaptation gain and the update rate for each Lagrange multiplier, respectively.
\color{red}
% The Eq.~\eqref{eq. adaptation law L max} is analog to using an active set $\mathcal I=\{j\ \vert \ c_j\ge 0\}$ in constrained optimization theory which excludes terms of satisfied constraints in adaptation law.
\color{black}

Using the chain rule, the gradient of the objective function with respect to the weights (\ie $\partial J/\partial {\hat{\mv\theta}}$) in \eqref{eq. adaptation law th} can be represented as
\begin{equation}
    {\partial J\over\partial {\hat{\mv\theta}}}
    =
    {\partial \mv\xi\over\partial {\hat{\mv\theta}}}^\top 
    \mm W
    {\partial \mv\xi\over\partial {\hat{\mv\theta}}}
    .
\end{equation}
The calculation of $\partial J/\partial {\hat{\mv\theta}}$ is not straightforward because of the dynamics of $\xi$.
Using the forward sensitivity method presented in \cite{Sengupta:2014aa}, the sensitivity equation can be obtained as follows:
\begin{equation}
    \dot {\mm \eta} =
    \mm A_\xi - \mm B_\xi 
    % (\partial \hat{\mv\Phi}/\partial {\hat{\mv\theta}})
    {\partial \hat{\mv\Phi}\over\partial {\hat{\mv\theta}}}
    ,
    \label{eq. sen eq}
\end{equation}
where ${\mm \eta}\triangleq \partial \mv\xi/\partial {\hat{\mv\theta}}\in\R^{2n\times \Xi}$ denotes the sensitivity of the weights to the augmented error.
The initial value of ${\mm \eta}$ is zero since the initial $\mv\xi$ is independent of the weights.
By decomposing for each layer, the dynamics of ${\mm \eta}_i\triangleq \partial \mv\xi/\partial {\hat{\mv\theta}}_i\in\R^{2n\times \Xi_i}$ can be represented as
\begin{equation}
    \begin{aligned}
        \dot{\mm \eta} =&
        \begin{bmatrix}
            {\mm \eta}_1&{\mm \eta}_0
        \end{bmatrix}'
        \\
        =&
        \mm A_\xi
        % \begin{bmatrix}
        [
            {\mm \eta}_1
            \ \ 
            {\mm \eta}_0
        ]
        % \end{bmatrix}
        -\mm B_\xi
        % \begin{bmatrix}
        [
            (\mm I_{l_2}\otimes \hat{\mv\phi}^\top )
            \ \ 
            \hat {\mm V}_1^\top \hat{\mv\phi}'(\mm I_{l_1}\otimes {\mv q_{NN}}^\top )
        ]
        % \end{bmatrix}
        .
    \end{aligned}
\end{equation}
The calculation of $\partial \hat{\mv\Phi}/\partial{\hat{\mv\theta}}$ is introduced in \cite{Patil:2022aa}.
In conclusion, the gradient of the objective with respect to the weights can be obtained as ${\partial J/\partial {\hat{\mv\theta}}} = {\mm \eta}^\top\mm W\mv\xi$ by simulating the sensitivity equation \eqref{eq. sen eq}.

On the other hand, the gradient of the constraints with respect to the weights can be represented as follows:
\begin{equation}
    \frac{\partial c_{\theta_0}}{\partial{\hat{\mv\theta}}}
    = 
    \begin{bmatrix}
        \mm 0_{\Xi_1\times 1} \\
        2{\hat{\mv\theta}}_0 
    \end{bmatrix}
    ,\quad
    \frac{\partial c_{\theta_1}}{\partial{\hat{\mv\theta}}}    
    = 
    \begin{bmatrix}
        2{\hat{\mv\theta}}_1 \\
        \mm 0_{\Xi_0\times 1}
    \end{bmatrix}
    .
\end{equation}

@@

\begin{theorem}
For the dynamical system in \eqref{eq. system dynamics 1}, the proposed controller \eqref{eq. control law} and the adaptation law \eqref{eq. adaptation law th}, \eqref{eq. adaptation law L} and \eqref{eq. adaptation law L max} ensure the boundedness of the tracking error $\mv \xi$ and the weight estimation ${\hat{\mv\theta}}$, provided that the control gains ${k_q}$ and ${k_z}$ satisfy \eqref{eq. ctrl stable condition}.
\end{theorem}

@@

\subsection*{Step 1: Boundedness of ${\hat{\mv\theta}}_1,{\mm \eta}_1,\mv \xi$}

@@

The dynamics of $\mv \xi$ can be represented as
\begin{equation}    
    \dot {\mv\xi} = \mm A_\xi\mv \xi +\mm B_\xi
    (
        -\hat {\mm V}_1^\top \hat{\mv\phi}+\mv w(t)
    )
\end{equation}
where $\mv w(t)\triangleq {\mm V}_1^{*T}\mv\phi^*+\mv \epsilon$ is a lumped residual term, which is bounded as $\Vert\mv  w(t)\Vert\le \bar w< 0$ since $\Vert\mv\theta_1^*\Vert$ and $\Vert\mv\phi^*\Vert$ are bounded.
The dynamics of ${\mm \eta}_1$ and ${\hat{\mv\theta}}_1$ are represented as
\begin{equation}
    \begin{aligned}
        \dot{\mm \eta}_1 =
        & 
        \mm A_\xi {\mm \eta}_1 -\mm B_\xi (\mm I_{l_{2}}\otimes \hat{\mv\phi}^\top )
        \\
        \dot{{\hat{\mv\theta}}}_1 =
        & -\alpha 
        (
            {\mm \eta}_1^\top \mm W\mv\xi+2\lambda_{\theta_1} {\hat{\mv\theta}}_1
        ).
    \end{aligned} 
\end{equation}
According to \cite[Chap.~4 T.~1.9]{Desoer:2009aa}, the boundedness of ${\mm \eta}_1$ can be obtained since $\mm A_\xi$ is stable and the residual term $-\mm B_\xi (\mm I_{l_{2}}\otimes \hat{\mv\phi}^\top )$ is bounded.

>>>>

\newcommand*{\q}{\mv{q}}
\newcommand*{\dq}{\dot{\q}}
\newcommand*{\ddq}{\ddot{\q}}
\newcommand*{\qd}{\mv{q_d}}
\newcommand*{\dqd}{\dot{\qd}}
\newcommand*{\ddqd}{\ddot{\qd}}

@@

\section*{Notation}
In this study, the following notation is used:

\begin{itemize}
    \item $\otimes$ denotes the Kronecker product \cite[Definition 7.1.2]{Bernstein:2009aa}.
    \item $\mv x=[x_i]_{i\in[1,\cdots,n]}\in\R^n$ denotes a vector.
    % \item $\mv x=[x_i]_{i\in[1,\cdots,n]}\in\R^n$ and $
    %     \mm A
    %     := 
    %     [a_{ij}]
    %     _{
    %         i\in[1,\cdots,n],j\in[1,\cdots ,]
    %     }\allowbreak\in\R^{n\times m}
    %     $ denotes a vector and a matrix.
    \item $\myrow_i(\mm A)$ denotes the $i\textsuperscript{th}$ row of the matrix $\mm A\in\R^{n\times m}$. 
    \item $\myvec(\mm A):= [\myrow_1(\mm A^\top)  ,\cdots,\myrow_m(\mm A^\top)  ]^\top   $ for $\mm A\in\R^{n\times m}$.
    % \item $\text{col}_i(A)$ and $\myrow_j(A)$ denote the $i\textsuperscript{th}$ column and $j\textsuperscript{th}$ row of $A\in\R^{n\times m}$, respectively.
    % \item $\myvec(A):= [\text{col}_1(A)^\top  ,\cdots,\text{col}_m(A)^\top  ]^\top   $ for $A\in\R^{n\times m}$.
    \item $\lambda_\text{min}(\mm A)$ denotes the minimum eigenvalue of the matrix $\mm A\in\R^{n\times n}$.
    \item $\mm I_n$ denotes the $n\times n$ identity matrix and $\mm 0_{n\times m}$ denotes the $n\times m$ zero matrix.
\end{itemize}

@@

@@

Consider an uncertain Euler‒Lagrange system modeled as
\begin{equation}
    \mm M\ddq + \mm V_m\dq + \mm F + \mm G + \mv\tau_d
    =
    \mv\tau 
    \label{eq. system dynamics 1}
\end{equation}
where $\q\in \mathbb{R}^n$ and $\mv \tau\in\R^n$ denote the generalized coordinate and the control input, respectively; $\mm M :=\mm M(\q)\in\mathbb{R}^{n\times n}$, $mm V_m:= \mm V_m(\q,\dq)\in\mathbb{R}^{n\times n}$, and $\mm G:=\mm G(\q)\in\mathbb{R}^{n}$ denote the unknown system function matrices; and $\mm F:=\mm F(\dq)\in\mathbb{R}^{n}$ denotes the external force.
Using the user-designed matrices $\mm M_0>0,\mm V_{m0}$ and $\mm G_0$, \eqref{eq. system dynamics 1} can be represented as 
\begin{equation}
    \mm M_0\ddq+\mm V_{m0}\dq+\mm G_0 = \mv \tau + \mv f(\q,\dq,\ddq)
    \label{eq. system dynamics 2}
\end{equation}
where $\mv f(\q,\dq,\ddq) := -(\mm M-\mm M_0)\ddq-(\mm V_m-\mm V_{m0})\dq -(\mm G-\mm G_0) -\mm F$ denotes the residual unknown term.

Hence, the objective of the control design is to make $\q$ track the continuously differentiable desired trajectory $\qd(t):\R\to \R^n$ under the unknown terms $\mv f$.

@@

The backstepping control-based approach is utilized to generate a reference signal $\mv z^*:= -{k_q}\tilde {\q}+\dq_d$ for $\mv {z}:= \dq$, where $\tilde {\q}:= \mv {q} - \qd$ and $k_q\in\R_{>0}$.
The desired stabilizing controller can be designed as follows:
\begin{equation}
    \mv \tau^* = 
    -\mm M_0\cdot ({k_z}\tilde {\mv z})
    + 
        ( 
            -\mm M_0\tilde {\q}+\mm V_{m0}{\mv z}+\mm G_0-\mv f+\mm M_0 \dot {\mv z}^*
        )
    \label{eq. desired control}
\end{equation}
where $\tilde {\mv z}:= \mv z-\mv z^*$ and $k_z\in\R_{>0}$.
Note that the desired controller cannot be realized because of $\mv f$.

To approximate the desired controller, an NN is utilized.
The NN with a single hidden layer is represented as 
\begin{equation}
    \NN({\q_{NN}};\wth):= \wV_1^\top \act(\wV_0^\top {\q_{NN}})
\end{equation}
where ${\q_{NN}}\in\R^{l_0+1}$ denotes the NN input vector, $\wV_i\in\R^{(l_i+1)\times l_{i+1}},\ i\in\{0,1\}$ denotes the weight matrix of the $i\textsuperscript{th}$ layer and $\act:\R^{l_1}\to\R^{l_1+1}$ denotes the activation function layer.
The activation function layer consists of an elementwise nonlinear function $\sigma(\cdot)$ and an augmented $1$ to combine the bias terms in the weight matrix (\ie $\act(\mv x) = \left(\sigma(\mv x_{(1)}),\cdots, \sigma(\mv x_{(l_1)}), 1\right)^\top $).
For further simplicity, let $\wth:=\left(\wth_1^\top ,\wth_0^\top \right)^\top \in\R^{\Xi}$ denote the total weight vector, where $\wth_i:= \myvec(\wV_i)\in\R^{\Xi_i}$ denotes the vectorized weights.
$\Xi_i=(l_i+1)\cdot l_{i+1}$ and $\Xi=\Xi_0+\Xi_1$ denote the number of each layer and total weights, respectively.

Using this NN, the desired controller $\mv \tau^*$ can be approximated by the ideal weight vector $\wth^*$ for a compact subset $\Omega_{NN}\in\R^{l_0+1}$ to $\epsilon$ accuracy such that $\sup_{{\q_{NN}}\in\Omega_{NN}} \allowbreak \norm{\NN({\q_{NN}};\wth^*) - \mv \tau^*}=\epsilon<\infty$ \cite{Cybenko:1989aa}.
The ideal weight vector $\wth^*$ is typically assumed to be bounded.
Then, using the estimated weight vector $\estwth=(\estwth_1^\top ,\estwth_0^\top )^\top $ of $\wth^*=(\wth_1^{*\top},\wth_0^{*\top})^\top $, the desired controller $\mv \tau^*\approx -\NN(\q_{NN};\wth^*)-\epsilon$ can be approximated as follows:
\begin{equation}
    \mv \tau := -\NN({\q_{NN}};\estwth)
    .
    \label{eq. control law}
\end{equation}
For further sections, let $\idealNN:=\NN({\q_{NN}};\idealwth)$ and $\NN^* :=\NN({\idealwV}_0^{\top}{\q_{NN}})$, and let $\estNN:={\NN}({\q_{NN}};\estwth)$, $\estNN :=\NN(\estwV_0^{\top}{\q_{NN}})$ and $\estNN' := \pptfrac{\estNN}{\estwV_0^\top \q_{NN}}$.

Using \eqref{eq. system dynamics 2}, \eqref{eq. desired control}, and \eqref{eq. control law}, the error dynamics can be obtained as the first-order system of augmented error $\mv \xi:=\left(\tilde {\q}^\top ,\tilde {\mv z}^\top\right)^\top \in\R^{2n}$ as follows:
\begin{equation}
    \dot{\mv \xi} 
    = 
    \mm A_\xi{\mv \xi}+\mm B_\xi(\NN^*-\estNN+\mv\epsilon)
\end{equation}
where $\mv \epsilon\in\R^2$ is bounded approximation error and 
\begin{equation}
    \mm A_\xi := 
    \begin{bmatrix}
        -{k_q} \mm I_{n} &\mm I_{n}
        \\
        -\mm I_{n}& -{k_z} \mm I_{n}
    \end{bmatrix}
    ,\ 
    \mm B_\xi := 
    \begin{bmatrix}
        \mm 0_{n\times n}\\\mm M_0^{-1}
    \end{bmatrix}.
\end{equation}

@@

For the boundedness of the weights, weight norm constraints are imposed on the adaptation process such that $c_{\theta_i}=\norm{{\estwth}_i}^2-\bar\theta^2_i,\ i\in[0,1]$.
The control problem can be reformulated into a constrained optimization problem as follows:
\begin{equation}
    \begin{matrix}
        \min_{{\estwth}} \ J(\mv \xi;{\estwth})
        :=
        {1\over 2}
        \mv \xi^\top \mm \Lambda \mv \xi
        \\ \\
        \begin{aligned}
        \text{subject to\;}&c_{j}\le0, \quad j\in\mathcal I 
        = 
        \{\theta_0,\theta_1\}
        ,
        \end{aligned}
    \end{matrix}
\end{equation}
where $\mm \Lambda=\mm \Lambda^\top >0$ denotes the weighting matrix.
In this optimization problem, $\mv \xi$ is considered a predefined parameter.
The corresponding Lagrangian function is defined as
\begin{equation}
    L(
        \mv \xi,
        {\estwth},
        [\lambda_j]_{j\in\mathcal I}
    )
    := 
    J(\mv \xi;{\estwth})
    +
    \textstyle\sum_{j\in\mathcal I} \lambda_jc_j({\estwth})
    .
\end{equation}
% where $\mathcal I=\{j\ \vert \ c_j\ge 0\}$ denotes the active set.
To solve the dual problem $\min_{{\estwth}} \max_{[\lambda_j]_{j\in\mathcal I}}  L$, the adaptation law is derived as follows:
\begin{subequations}
    \begin{align}
            \ddtt {{\estwth}}
            &
            =
            -\alpha 
            \pptfrac{L}{{\estwth}}
            =-\alpha 
            \left(
                \pptfrac{J}{{\estwth}}
                +
                \textstyle\sum_{j\in\mathcal{I}}
                \lambda_j 
                \pptfrac{c_j}{\estwth}
            \right),
        \label{eq. adaptation law th}
            \\
            \ddtt\lambda_j
            & 
            = 
            \beta_j
            \pptfrac{L}{\lambda_j} 
            = 
            \beta_j c_j ,
            \quad\quad\quad\quad      \      
            \forall j\in\mathcal I,
        \label{eq. adaptation law L}
            \\
            \lambda_j 
            & 
            = 
            \max(\lambda_j,0) ,
            \quad\quad\quad\quad\ \ \ \ \ 
            \forall j\in\mathcal I,
        \label{eq. adaptation law L max}
    \end{align}
    \label{eq. adaptation laws}
\end{subequations}
where $\alpha\in\R_{>0}$ and $\beta_j\in\R_{>0}$ denote the adaptation gain and the update rate for each Lagrange multiplier, respectively.
\color{red}
% The Eq.~\eqref{eq. adaptation law L max} is analog to using an active set $\mathcal I=\{j\ \vert \ c_j\ge 0\}$ in constrained optimization theory which excludes terms of satisfied constraints in adaptation law.
\color{black}

Using the chain rule, the gradient of the objective function with respect to the weights (\ie $\pptfrac{J}{\estwth}$) in \eqref{eq. adaptation law th} can be represented as
\begin{equation}
    \pptfrac{J}{\estwth}
    =
    \pptfrac{\mv\xi}{\estwth}^\top 
    \mm \Lambda
    \pptfrac{\mv\xi}{\estwth}
    .
\end{equation}
The calculation of $\pptfrac{J}{\estwth}$ is not straightforward because of the dynamics of $\mv\xi$.
Using the forward sensitivity method presented in \cite{Sengupta:2014aa}, the sensitivity equation can be obtained as follows:
\begin{equation}
    \ddtt{\mm \eta} =
    \mm A_\xi - \mm B_\xi 
    % (\partial \estNN/\partial {\estwth})
    \pptfrac{\estNN}{\estwth}
    ,
    \label{eq. sen eq}
\end{equation}
where ${\mm \eta}:= \pptfrac{\mv\xi}{\estwth}\in\R^{2n\times \Xi}$ denotes the sensitivity of the weights to the augmented error.
The initial value of ${\mm \eta}$ is zero since the initial $\mv\xi$ is independent of the weights.
By decomposing for each layer, the dynamics of ${\mm \eta}_i:= \pptfrac{\mv\xi}{\estwth}_i\in\R^{2n\times \Xi_i}$ can be represented as
\begin{equation}
    \begin{aligned}
        \ddtt{\mm \eta} 
        \!=\!&
        \begin{bmatrix}
            {\mm \eta}_1&{\mm \eta}_0
        \end{bmatrix}'
        \\
        \!=\!&
        \mm A_\xi
        % \begin{bmatrix}
        [
            {\mm \eta}_1
            \ \ 
            {\mm \eta}_0
        ]
        % \end{bmatrix}
        \!-\!
        \mm B_\xi
        % \begin{bmatrix}
        [
            (\mm I_{l_2}\!\otimes\!\estact^\top )
            \ \ 
            \estwV_1^\top \estact'(\mm I_{l_1}\otimes {\q_{NN}}^\top )
        ]
        % \end{bmatrix}
        .
    \end{aligned}
\end{equation}
The calculation of $\pptfrac{\estNN}{\estwth}$ is introduced in \cite{Patil:2022aa}.
In conclusion, the gradient of the objective with respect to the weights can be obtained as $\pptfrac{J}{\estwth} = {\mm \eta}^\top\mm \Lambda\mv\xi$ by simulating the sensitivity equation \eqref{eq. sen eq}.

On the other hand, the gradient of the constraints with respect to the weights can be represented as follows:
\begin{equation}
    \pptfrac{c_{\theta_0}}{{\estwth}}
    = 
    \begin{bmatrix}
        \mm 0_{\Xi_1\times 1} \\
        2{\estwth}_0 
    \end{bmatrix}
    ,\quad
    \pptfrac{c_{\theta_1}}{{\estwth}}    
    = 
    \begin{bmatrix}
        2{\estwth}_1 \\
        \mm 0_{\Xi_0\times 1}
    \end{bmatrix}
    .
\end{equation}

@@

\begin{theorem}
For the dynamical system in \eqref{eq. system dynamics 1}, the proposed controller \eqref{eq. control law} and the adaptation law \eqref{eq. adaptation laws} ensure the boundedness of the tracking error $\mv \xi$ and the weight estimation ${\estwth}$, provided that the control gains ${k_q}$ and ${k_z}$ satisfy \eqref{eq. ctrl stable condition}.
\end{theorem}

@@

\subsection*{Step 1: Boundedness of ${\estwth}_1,{\mm \eta}_1,\mv \xi$}

@@

The dynamics of $\mv \xi$ can be represented as
\begin{equation}    
    \ddtt {\mv\xi} = \mm A_\xi\mv \xi +\mm B_\xi
    (
        -\estwV_1^\top \estact+\mv w(t)
    )
\end{equation}
where $\mv w(t):= {\idealwV}_1^{\top}\idealact+\mv \epsilon$ is a lumped residual term, which is bounded as $\norm{\mv  w(t)}\le \bar w< 0$ since $\norm{\idealwth_1}$ and $\norm{\idealact}$ are bounded.
The dynamics of ${\mm \eta}_1$ and ${\estwth}_1$ are represented as
\begin{equation}
    \begin{aligned}
        \ddtt{\mm \eta}_1 =
        & 
        \mm A_\xi {\mm \eta}_1 -\mm B_\xi (\mm I_{l_{2}}\otimes \estact^\top )
        ,
        \\
        \ddtt{{\estwth}}_1 =
        & -\alpha 
        (
            {\mm \eta}_1^\top 
            \mm \Lambda
            \mv\xi
            +
            2\lambda_{\theta_1} {\estwth}_1
        )
        .
    \end{aligned} 
\end{equation}
According to \cite[Chap.~4 T.~1.9]{Desoer:2009aa}, the boundedness of ${\mm \eta}_1$ can be obtained since $\mm A_\xi$ is stable and the residual term $-\mm B_\xi (\mm I_{l_{2}}\otimes \estact^\top )$ is bounded.

Define the Lyapunov function $V_1=\tfrac{1}{2}\mv\xi^\top \mm P\mv\xi+\tfrac{1}{2\alpha}{\estwth}_1^\top {\estwth}_1$, with the Lyapunov equation $\mm A_\xi^\top \mm P+\mm P\mm A_\xi=-\mm Q$, where $\mm A_\xi<0,\mm P=\mm P^\top >0$, and $\mm Q>0$.
Using a proposition $\estwV_1^\top \estact = \myvec(\estwV_1^\top \estact)=\myvec(\estact^\top \estwV_1) = (\mm I_{l_2}\otimes {\estact}^\top )\myvec(\estwV_1)=(\mm I_{l_2}\otimes \estact^\top ){\estwth}_1$ \cite[Proposition~(7.1.9)]{Bernstein:2009aa}, the time derivative of $V_1$ is
\begin{equation}
    \begin{aligned}
        \ddtt \! {V}_1 
        \!=\!& 
        \tfrac{1}{2}
        \mv \xi^\top 
            (
                \mm A_\xi^\top \mm P+\mm P\mm A_\xi
            )
        \mv \xi
        \!+\!
        \mv \xi^\top \mm P (
            \!-\!\mm B_\xi\estwV_1^\top \estact +\mm B_\xi \mv w(t) 
        )\\
        &
        +{\estwth}_1^\top 
        \!
        (\!
            -{\mm \eta}_1^\top \mm \Lambda \xi-2\lambda_{\theta_1} {\estwth}_1 
        \!)
        \\
        =& 
        -\tfrac{1}{2}
        \mv \xi^\top \mm Q\mv \xi 
        -
        \mv \xi^\top \mm P\mm B_\xi(\mm I_{l_2}\otimes \estact^\top ){\estwth}_1 
        +\mv \xi^\top \mv \Delta
        \\
        &
        -{\estwth}_1^\top {\mm \eta}_1^\top \mm \Lambda\mv \xi
        -2\lambda_{\theta_1} {\estwth}_1^\top {\estwth}_1
        \\
        \le&
        -
        \tfrac{1}{2}
        \lambda_\text{min}(\mm Q)\norm{\mv \xi} ^2
        +\bar\Delta \norm{\mv \xi}
        +\bar M\norm{\mv \xi} \norm{{\estwth}_1}
        \\
        &
        -2\lambda_{\theta_1}
        \norm{{\estwth}_1}^2
        \\
        \le& 
        (
        -{\lambda_\text{min}(\mm Q)/ 2} +{\tfrac{\bar M}{2}}
        )
        \norm{\mv \xi}^2 +\bar\Delta \norm{\mv \xi}
        \\
        &+ 
        (
        -2\lambda_{\theta_1} 
        +\tfrac{\bar M}{2}
        )
        \norm{\tilde{\wth}_1}^2 
        \end{aligned}
        \label{eq. lya 1}
\end{equation}
where $\mv \Delta:= \mm P\mm B_\xi \mv w(t)$ and $\mm M:= -\mm P\mm B_\xi(\mm I_{l_2}\otimes \estNN^\top )+\mm \Lambda{\mm \eta}_1$ are bounded such that $\norm{\mv \Delta}\le\bar\Delta<\infty$ and $\norm{\mm M}_F\le \bar M< \infty$, respectively.
